Just Add IFC! A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-03-02 paper/intro.tex

One alternative is a \textit{coarse-grained} approach to IFC.  Here,
computational units are associated with a floating label which protects
all data in this unit.  A recent system named SWAPI adds IFC to the
browser by identifying \emph{web workers} as an existing computational
unit.  Communication between workers can efficiently be mediated in an
IFC aware way to ensure non-interference, where sensitive workers cannot
influence the computations in less sensitive workers.

This coarse-grained approach as employed by SWAPI provides a number
of advantages; (1) reduced runtime overhead because checks need only
be performed at isolation boundaries, (2) minimal changes to an
existing programming language when adding IFC, which (3) allows
the reuse of existing programs.  Finally, (4) the simplicity
of a single security label per computational unit facilitates
understanding and easy reasoning about the correctness of the
system, in particular without having to reason about all the
technical details of the semantics of the programming language
at hand.

The recognition of the simplicity of this coarse-grained approach
to IFC allows us to generalize it to a large class of programming
languages.  In this paper we identify the key concepts of coarse-grained
dynamic IFC and show how these can be combined with a programming
language, which we will call the target language.  We achieve this
by using an embedding that combines the operational semantics
of the core IFC calculus and the target language.
At the core of our IFC system we use a sandboxing operation that allows
the execution of potentially sensitive computations in isolation,
while communicating through clearly specified interfaces with other
sandboxed operations.
We formally state and prove the non-interference guarantees that our
system gives, independently of the choice of the target language.

One challenge in the design of such a general IFC calculus that
is supposed to work for a large class of programming languages is
that the stated semantics might not correspond well to the
semantics of a concrete implementation for a particular choice
of target language.  We address this issue by using the notation
of isomorphism to characterize a condition that guarantees
non-interference of the concrete system without having to prove
this property manually.

Our key contributions can be summarized as follows:
\begin{itemize}
  \item We give formal semantics for a core IFC language that
  captures the essence of coarse-grained information flow control.
  \item We show how a large class of target languages can be combined
  with this IFC language and prove that the resulting system provides
  non-interference.
%  \item We show how to extend the Matthews-Findler method from just
%  languages that are not simply expressions evaluating to values,
%  but may be collections of threads executing nondeterministically.
  \item We provide a proof technique to show the non-interference
  of a concrete semantics for a potentially optimized IFC language
  by means of an isomorphism, and show a class of restrictions on
  the IFC calculus that preserves non-interference.
  \item We sketch how the core IFC language can be enriched with
  more advanced concepts such as labeled values, privileges or a
  notation of clearance.
  \item We connect our formal semantics to real implementations of
  coarse-grained IFC systems for Javascript and Haskell, and explain
  how our system could be used to enforce IFC in C.
  \item We informally connect our results to a monadic interpretation
  of IFC which guided our design.
\end{itemize}

>>>>

One alternative is a \textit{coarse-grained} approach to IFC.  Here,
computational units are associated with a floating label which protects
all data in this unit.  A recent system named SWAPI adds IFC to the
browser by identifying \emph{web workers} as an existing computational
unit.  Communication between workers can efficiently be mediated in an
IFC aware way to ensure non-interference, where sensitive workers cannot
influence the computations in less sensitive workers.
This coarse-grained approach provides a number
of advantages: (1) reduced runtime overhead because checks need only
be performed at isolation boundaries, (2) minimal changes to an
existing programming language when adding IFC, which (3) allows
the reuse of existing programs.  Finally, (4) the simplicity
of a single security label per computational unit facilitates
understanding and easy reasoning about the correctness of the
system, in particular without having to reason about all the
technical details of the semantics of the programming language
at hand.

In fact, the simplicity of this coarse-grained approach
to IFC allows us to generalize it to a large class of programming
languages.  In this paper, we identify the key concepts of coarse-grained
dynamic IFC and show how these can be combined with \emph{any} programming
language.  We achieve this
by using an embedding that combines the operational semantics
of the core IFC calculus and the target language.
At the core of our IFC system we use a sandboxing operation that allows
the execution of potentially sensitive computations in isolation,
while communicating through clearly specified interfaces with other
sandboxed operations.
We formally state and prove the non-interference guarantees that our
system gives, independently of the choice of the target language.

One challenge in the design of our system is that the semantics for the
retrofitted IFC language may not reflect the actual implementation in a
particular choice of target language.  We partially address this issue
by characterizing isomorphisms between the operational semantics of our
defined language and a concrete implementation, showing that if this
relationship holds, then noninterference in the abstract specification
carries over to the concrete specification.

Our key contributions can be summarized as follows:
\begin{itemize}
  \item We give formal semantics for a core IFC language that
      captures the essence of coarse-grained information flow control.
  We then show how a large class of target languages can be combined
  with this IFC language and prove that the resulting system provides
  non-interference. (Section~\ref{sec:retrofit})
%  \item We show how to extend the Matthews-Findler method from just
%  languages that are not simply expressions evaluating to values,
%  but may be collections of threads executing nondeterministically.
  \item We provide a proof technique to show the non-interference
  of a concrete semantics for a potentially optimized IFC language
  by means of an isomorphism, and show a class of restrictions on
  the IFC calculus that preserves non-interference. (Section~\ref{sec:formal})
  \item We sketch how the core IFC language can be enriched with
  more advanced concepts such as labeled values, privileges or a
  notation of clearance. (Section~\ref{sec:extensions})
  \item We connect our formal semantics to real implementations of
  coarse-grained IFC systems for Javascript and Haskell, and explain
  how our system could be used to enforce IFC in C. (Section~\ref{sec:real})
  \item We informally connect our results to a monadic interpretation
      of IFC which guided our design. (Section~\ref{sec:monad})
\end{itemize}
